---
title: "ALTa Ciencia Abierta 1 2024"
format:
 gfm:
    fig-width: 7
    fig-height: 5
    toc: true
execute: 
  echo: false
  warning: false
---

```{r}
# Paquetes


# Manipulación de datos y visualización
library(tidyverse)
library(ggthemes)
library(ggstats)
library(glue)
library(knitr)
library(kableExtra)

# Mapas
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(countrycode)
library(ggiraph)

# devtools::install_github("yutannihilation/ggsflabel")
library(ggsflabel)

# Importar datos
library(readxl)

# Tamaño base de la letra de gráficos
base_size = 16




```

```{r}
#| label: setup-globals
#| include: false

source("../_intervalo_confianza.R")
```

```{r}

# Datos del formulario de inscripción y el pre test de investigación

inscripciones <- read_csv("data/inscripciones_total.csv")
participantes <- read_csv("data/inscripciones_participantes.csv")
investigacion <- read_csv("data/inscripciones_investigacion.csv")
accesibilidad <- read_csv("data/accesibilidad.csv")
grupos <- read_csv("data/grupos_subrepresentados.csv")

# Formularios de evaluación

evaluaciones <- read_csv("data/Evaluaciones.csv")

# Datos de la encuesta final y el post test de investigacion

encuesta_final <- read_csv("data/encuesta_final.csv")

# Datos de slack

slack <- read_csv("data/slack_inscriptos.csv")


# Datos de zoom

zoom_tiempo <- read_csv("data/zoom_tiempo.csv")
zoom_final <- read_csv("data/zoom_final.csv")






```

## Meet the attendees!

```{r}
# Tabla con porcentaje de inscriptos por país
paises <- inscripciones %>%
  group_by(pais) %>%
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(porcentaje = round(n/sum(n)*100,1)) %>% 
  arrange(desc(porcentaje))

paises <- paises %>% 
  mutate(pais = str_replace(pais, pattern = "República Checa", replacement = "Chequia"))

# Vector con los distintos países de los participantes
paises_metadocencia <- paises %>%
  drop_na() %>%
  pull(pais)

# Información de longitud y latitud para crear el mapa
mapamundi <- ne_countries(scale = "medium", returnclass = "sf") 

# Códigos iso3 para cada país, y nombre en español e inglés
codigos <- countrycode::codelist %>% 
  select(pais_ES = cldr.name.es, iso_a3 = iso3c) %>% 
  drop_na(pais_ES) %>% 
  filter(iso_a3 %in% mapamundi$iso_a3)

# Unir nombres en español al mapamundi
mapamundi <- mapamundi %>% 
  left_join(codigos, by = "iso_a3") 

# Sumar información sobre los inscriptos al mapamundi
mapamundi <- mapamundi %>% 
  mutate(metadocencia = ifelse(pais_ES %in% paises_metadocencia, TRUE, FALSE)) %>% 
  left_join(paises, by = c("pais_ES" = "pais")) %>% 
  mutate(etiqueta = ifelse(metadocencia == TRUE, 
                           str_c(name, " ", porcentaje, "%"), NA))


paises_in <- data.frame(MD = mapamundi$metadocencia,
                        pais_in = mapamundi$name_long,
                        pais_es = mapamundi$pais_ES,
                        n = mapamundi$n,
                        porcentaje = mapamundi$porcentaje)

# comentar este código si se quieren los datos en español
paises_in <- paises_in %>%
  filter(MD) %>%
  select(pais_in, n, porcentaje) %>%
  arrange(desc(n))

paises <- paises_in

```

### Where do they come from? (n = `r sum(paises$n)`, `r sprintf('%.1f', sum(paises$n)/nrow(inscripciones)*100)`%)

A total of `r nrow(inscripciones)` participants signed up from `r nrow(paises)` countries, `r paises[1,3]`% (n = `r paises[1,2]`) from `r paises[1,1]`, followed by `r paises[2,3]`% (n = `r paises[2,2]`) from `r paises[2,1]` and `r paises[3,3]`% (n = `r paises[3,2]`) from `r paises[3,1]`. The remaining `r sum(paises[4:nrow(paises),3])`% (n = `r sum(paises[4:nrow(paises),2])`) were from `r str_c(paises$pais[4:nrow(paises)], collapse = ", ")`.

#### Figure 1. Country of residence of the people that signed up

```{r}
# Visualización de mapa
mapamundi %>% 
  ggplot() +
  geom_sf(aes(fill = metadocencia), color = "white", size = .3) +
  labs(title = "",
       subtitle = "",
       x = "",
       y = "") +
  scale_fill_manual(values = c("gray", "#c83737")) +
  theme_map(base_size = base_size) +
  coord_sf(xlim = c(-160, 170), ylim = c(-55, 90)) +
  geom_sf_label_repel(aes(label = etiqueta),
                      force = 30, seed = 10,
                      direction = "both") +
    theme(
    panel.spacing = unit(c(-1, 0, -1, 0), "cm"),
    plot.margin = unit(c(-0.3, 0, -0.15, 0), "null"), # Adjust margins to trim the plot
    legend.position = 'none'
  )






```

#### Figure 2. Country of residence of the people that signed up

```{r}
# Visualización de mapa
mapamundi %>% 
  ggplot() +
  geom_sf(aes(fill = metadocencia), color = "white", size = .3) +
  labs(title = "",
       subtitle = "",
       x = "",
       y = "") +
  scale_fill_manual(values = c("gray", "#c83737")) +
  theme_map(base_size = base_size) +
  coord_sf(xlim = c(-160, 170), ylim = c(-55, 90)) +
    theme(
    panel.spacing = unit(c(-1, 0, -1, 0), "cm"),
    plot.margin = unit(c(-0.3, 0, -0.15, 0), "null"), # Adjust margins to trim the plot
    legend.position = 'none'
  )



```

#### Figure 3. Country of residence of the people that signed up

```{r}

mapamundi %>% 
  ggplot() +
  geom_sf(aes(fill = n), color = "white", size = .3) +
  labs(title = "",
       subtitle = "",
       x = "",
       y = "") +
  theme_map(base_size = base_size) +
  coord_sf(xlim = c(-160, 170), ylim = c(-55, 90)) +
  theme(
    panel.spacing = unit(c(-1, 0, -1, 0), "cm"),
    plot.margin = unit(c(-0.3, 0, -0.15, 0), "null"),
    legend.position = "bottom"
  ) +
  scale_fill_gradient(low = "#E09090", high = "#c83737", na.value = "gray", name = "Participants", guide = guide_colorbar(title.position = "top")) +
  theme(legend.text = element_text(size = base_size*.6),
        legend.title = element_text(size = base_size*.8))
```

```{r}

pronom <- participantes %>% 
  filter(!is.na(pronombre)) %>% 
  count(pronombre) %>%
  mutate(pronombre = ifelse(pronombre == "El", "He|his", "She|her")) %>%
  mutate(
    por = round(n/ sum(n) * 100,1),
    etiqueta = paste0(pronombre, " (", por, "%)"))
```

::: {layout="[0.5, 0.5]"}
::: {#firstcol}

#### Figure 4. Pronouns used by the people that signed up
```{r}

# Hole size
hsize <- 2

pronom <- pronom %>% 
  mutate(x = hsize)

pronom %>%
  ggplot(aes(x = hsize, y = por, fill = pronombre)) +
  geom_col() +
  geom_label(aes(label = paste(pronombre, "\n", por, "%")),
             position = position_stack(vjust = 0.5), color = "white",
             label.size = 1,
             size = base_size*.5) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("#00506fff", "#c63838ff")) +
  xlim(c(0.2, hsize + 0.5)) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none")


```
:::

::: {#secondcol}
### Which pronouns do they use? (n = `r sum(pronom$n)`, `r round(sum(pronom$n)/nrow(inscripciones)*100,1)`%)

Out of the `r sum(pronom$n)` participants, `r pronom$por[1]`% (n = `r pronom$n[1]`) use the pronouns `r pronom$pronombre[1]`.
:::
:::

```{r}
equipo <- participantes %>% 
  filter(!is.na(equipo)) %>% 
  count(equipo) %>% 
  mutate(equipo = str_extract(equipo, "Sí|No")) %>%
  mutate(equipo = ifelse(equipo == "No", "No", "Yes")) %>%
  mutate(
    por = round(n/ sum(n) * 100,1),
    etiqueta = paste0(equipo, " (", por, "%)"))
```

::: {layout="[0.5, 0.5]"}
::: {#firstcol}

#### Figure 5. People that signed up as part of a team. 

```{r}

equipo <- equipo %>% 
  mutate(x = hsize)

equipo %>%
  ggplot(aes(x = hsize, y = por, fill = equipo)) +
  geom_col() +
  geom_label(aes(label = paste(equipo, "\n", por, "%")),
             position = position_stack(vjust = 0.5), color = "white",
             label.size = 1,
             size = base_size*.5) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("#00506fff", "#c63838ff")) +
  xlim(c(0.2, hsize + 0.5)) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none")

```
:::

::: {#secondcol}
### Are they attending as part of a team? (n = `r sum(equipo$n)`, `r round(sum(equipo$n)/nrow(inscripciones)*100,1)`%)

Out of the `r sum(equipo$n)` participants who answered this question, `r equipo$por[2]`% (n = `r equipo$n[2]`) joined as part of a team.
:::
:::

```{r}
n_educ <- participantes %>% 
  filter(!is.na(nivel_educativo)) %>% 
  count(nivel_educativo) %>% 
  mutate(por = round(n/sum(n)*100,1)) %>%
   mutate(nivel_educativo = fct_relevel(nivel_educativo, "Secundario completo", "Universitario/terciario incompleto", "Universitario/terciario completo", "Posgrado incompleto", "Posgrado completo")) %>% 
  mutate(nivel_educativo = fct_recode(nivel_educativo, "Completed high school"  = "Secundario completo", "In progress bachelor's degree" = "Universitario/terciario incompleto", "Completed bachelor's degree" = "Universitario/terciario completo", "Some graduate courses" = "Posgrado incompleto", "Graduate degree" = "Posgrado completo")) %>%
  arrange(nivel_educativo)

plot.height = n_distinct(n_educ$nivel_educativo)*0.5+0.5
```

### What is their educational level? (n = `r sum(n_educ$n)`, `r round(sum(n_educ$n)/nrow(inscripciones)*100,1)`%)

Almost two-thirds of the participants (`r sum(n_educ$por[3:5])`%, n = `r sum(n_educ$n[3:5])`) of the participants have an undergraduate degree, `r n_educ$por[5]`% (n = `r n_educ$n[5]`) has additionally completed postgraduate courses. Finally, `r sum(n_educ$por[1])`% (n = `r sum(n_educ$n[1])`) completed high school and `r sum(n_educ$por[2])`% (n = `r sum(n_educ$n[2])`) has taken or is currently taking university courses.

#### Figure 6. Educational level of the people that signed up

```{r, fig.width=7, fig.height=plot.height}

n_educ %>%
  ggplot(aes(nivel_educativo, por)) +
  geom_bar(stat = "identity",fill="#c83737") +
  ggtitle("")+
  xlab("") +
  ylab("Percentage (%)") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0)) +
  theme_minimal(base_size = base_size)

```

```{r}
formacion <- participantes %>% 
  filter(!is.na(area_formacion)) %>% 
  count(area_formacion) %>% 
  mutate(por = round(n/sum(n)*100,1)) %>%
    mutate(area_formacion = fct_reorder(area_formacion,n)) %>% 
  arrange(desc(por))

plot.height = n_distinct(formacion$area_formacion)*0.5+0.5
```

### What is their field of study? (n = `r sum(formacion$n)`, `r round(sum(formacion$n)/nrow(inscripciones)*100,1)`%)

We have attendees from `r nrow(formacion)` different fields of study. The `r formacion$por[1]`% (n = `r formacion$n[1]`) corresponds to `r formacion$area_formacion[1]`, followed by a `r formacion$por[2]`% (n = `r formacion$n[2]`) corresponding to `r formacion$area_formacion[2]` and a `r formacion$por[3]`% (n = `r formacion$n[3]`) to `r formacion$area_formacion[3]`. The remaining `r sum(formacion$por[4:nrow(formacion)])`% (n = `r sum(formacion$n[4:nrow(formacion)])`) are people of the following fields: `r str_c(formacion$area_formacion[4:nrow(formacion)], collapse = ", ")`.

#### Figure 7. Field of study of the people that signed up


```{r, fig.width=7, fig.height=plot.height}
formacion %>%
  ggplot(aes(area_formacion, por)) +
  geom_bar(stat = "identity",fill="#c83737") +
  ggtitle("")+
  xlab("") +
  ylab("Percentage (%)") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0)) +
  theme_minimal(base_size = base_size)


```

```{r}

invest <- participantes %>% 
    filter(!is.na(investigacion)) %>%
  count(investigacion) %>% 
  mutate(investigacion = ifelse(investigacion == "No", "No", "Yes")) %>%
  mutate(
    por = round(n/ sum(n) * 100,1),
    etiqueta = paste0(investigacion, " (", por, "%)"))

max <- participantes %>%
  filter(!is.na(investigacion_anios), investigacion_anios >= 20) %>%
  count()

anios <- IC_funcion(data = participantes, variable = "investigacion_anios")

```

### Do they participate in research projects? (n = `r nrow(investigacion)`, `r round(nrow(investigacion)/nrow(inscripciones)*100,1)`%)


Out of the people who signed up, `r invest$por[2]`% (n = `r invest$n[2]`) participates in research projects. Of those participants, 50% (n = `r round(nrow(investigacion)/2,0)`) have between 0 and `r median(participantes$investigacion_anios, na.rm = TRUE)` years of experience. Those who participate in research (`r invest$por[2]`%, n = `r invest$n[2]`) have an average of `r anios$media` (sd = `r anios$DE`, IC = `r anios$IC`) years of experience. A small group of them (`r round(max$n[1]/invest$n[2]*100,1)`%, n = `r max$n[1]`) has more than 20 years of experience.

::: {layout="[0.5, 0.5]"}
::: {#firstcol}

#### Figure 8. Percentage of the people that signed up that participate in research projects. 

```{r}

invest <- invest %>% 
  mutate(x = hsize)

invest %>%
  ggplot(aes(x = hsize, y = por, fill = investigacion)) +
  geom_col() +
  geom_label(aes(label = paste(investigacion, "\n", por, "%")),
             position = position_stack(vjust = 0.5), 
             color = "white",
             label.size = 1,
             size = base_size*.5) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("#00506fff", "#c63838ff")) +
  xlim(c(0.2, hsize + 0.5)) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none")

```
:::

::: {#secondcol}

#### Figure 9. Years of experience in research for the people that participate in a research project. 

```{r}

participantes %>%
  filter(!is.na(investigacion_anios)) %>%
  ggplot(aes(x = investigacion_anios)) +
  geom_histogram(fill = "#C63838", color = "white") +
  theme_minimal(base_size = base_size) +
  labs(x = "Years of experience", y = "Count") +
  theme(axis.text = element_text(size = 20),
        axis.title = element_text(size = 30))



```
:::
:::

```{r}


plot.height = n_distinct(grupos$comunidad)*0.5+0.5
  
```

### Are they members of groups underrepresented in science? (n = `r sum(grupos$n)`, `r round(sum(grupos$n)/nrow(inscripciones)*100,1)`%)

Among the `r sum(grupos$n)` persons (`r round(sum(grupos$n)/nrow(inscripciones)*100,1)`%) who answered this question, `r grupos$por[1]`% (n = `r grupos$n[1]`) belongs to the category `r str_to_lower(grupos$comunidad[1])`, the `r grupos$por[2]`% (n = `r grupos$n[2]`) belongs to `r str_to_lower(grupos$comunidad[2])` and the `r grupos$por[3]`% (n = `r grupos$n[3]`) to `r str_to_lower(grupos$comunidad[3])`. The remaining `r sum(grupos$por[4:nrow(grupos)])`% (n = `r sum(grupos$n[4:nrow(grupos)])`) identify as belonging to the following groups: `r str_to_lower(str_c(grupos$comunidad[4:nrow(grupos)], collapse = ", "))`.

#### Figure 10. Underrepresented groups in technical and scientific contexts that participants identify with. 

```{r, fig.width=7, fig.height=plot.height}
grupos %>%
  ggplot(aes(fct_reorder(comunidad, por), por)) +
  geom_bar(stat = "identity",fill="#c83737") +
  ggtitle("")+
  xlab("") +
  ylab("Percentage (%)") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0)) +
  theme_minimal(base_size = base_size) +
  labs(caption = "More than one option can be selected")
```



The total percentage exceeds 100% because participants were asked to choose all the categories that could represent their intersectionalities of underrepresentation in science.


## Accessibility

```{r}


  


plot.height = n_distinct(accesibilidad$accesibilidad)*0.5+0.5

```

### What are their accessibility requirements? (n = `r sum(accesibilidad$n)`, `r round(sum(accesibilidad$n)/nrow(inscripciones)*100,1)`%)

This question was completed by `r sum(accesibilidad$n)` persons, a `r round(sum(accesibilidad$n)/nrow(inscripciones)*100,1)`% of the total of participants. Among those who answered, the biggest request is `r accesibilidad$accesibilidad_min[1]` (`r sprintf('%.1f', accesibilidad$por[1])`%, n = `r accesibilidad$n[1]`), followed by `r  accesibilidad$accesibilidad_min[2]` (`r sprintf('%.1f', accesibilidad$por[2])`%, n = `r accesibilidad$n[2]`). The remaining `r sprintf('%.1f', sum(accesibilidad$por[4:nrow(accesibilidad)]))`% (n = `r sum(accesibilidad$n[4:nrow(accesibilidad)])`) corresponds to people that requested: `r str_c(accesibilidad$accesibilidad_min[4:nrow(accesibilidad)], collapse = ", ")`.

#### Figure 11. Accessibility requirements by participants. 


```{r, fig.width=7, fig.height=plot.height}
accesibilidad  %>% 
  ggplot(aes(accesibilidad, por)) +
  geom_bar(stat = "identity",fill="#c83737") +
  ggtitle("")+
  xlab("") +
  ylab("Percentage (%)") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0)) +
  theme_minimal(base_size = base_size)
```


The total percentage exceeds 100% because participants were asked to choose all the categories that could represent their intersectionalities of underrepresentation in science.

## Background in Open Science

```{r}
publicacion <- investigacion %>% 
  select(conoces_proyecto:conoces_resultados2) %>% 
   mutate_all(~ ifelse(. == "No aplica/No sé/Prefiero no contestar", NA, .))

publicacion <- publicacion[rowSums(is.na(publicacion)) != ncol(publicacion), ]


```

### What is their knowledge about open publishing? (n = `r nrow(publicacion)`, `r round(nrow(publicacion)/nrow(inscripciones)*100,1)`%)


```{r}

## Español
# publicacion2 <- publicacion %>% 
#   pivot_longer(cols = everything(),
#                names_to = "Medida",
#                values_to = "Valor") %>% 
#   mutate(Medida = fct_recode(Medida, `Proyecto de investigación` = "conoces_proyecto",
# `Plan de gestión de datos` = "conoces_pgd",
# `Hipótesis y plan de análisis` = "conoces_hipotesis",
# `Datos recolectados` = "conoces_datos",
# `Código` = "conoces_codigo",
# `Instrumentos` = "conoces_instrumentos",
# `Resultados en revistas` = "conoces_resultados1",
# `Resultados en medios alternativos` = "conoces_resultados2")) %>% 
#   mutate(Valor = fct_relevel(Valor, "No conozco esta práctica",
#                              "Conozco la práctica pero nunca la he aplicado",
#                              "He aplicado la práctica")) %>% 
#   filter(!is.na(Valor))

# Inglés
publicacion2 <- publicacion %>% 
  pivot_longer(cols = everything(),
               names_to = "Medida",
               values_to = "Valor") %>% 
  mutate(Medida = fct_recode(Medida, `Pre-register a research project` = "conoces_proyecto",
`Data management plan` = "conoces_pgd",
`Hypothesis and analysis plan` = "conoces_hipotesis",
`Data collected` = "conoces_datos",
`Code` = "conoces_codigo",
`Instruments` = "conoces_instrumentos",
`Results in Open Access journals` = "conoces_resultados1",
`Results in alternative media` = "conoces_resultados2")) %>% 
  mutate(Valor = fct_relevel(Valor, "No conozco esta práctica",
                             "Conozco la práctica pero nunca la he aplicado",
                             "He aplicado la práctica"),
         Valor = fct_recode(Valor, 
                            "I don't know this practice" = "No conozco esta práctica",
  "I know this practice but I have never applied it" = "Conozco la práctica pero nunca la he aplicado",
  "I have applied this practice" = "He aplicado la práctica")) %>% 
  filter(!is.na(Valor))




porcentaje <- publicacion2 %>% 
              group_by(Medida, Valor) %>%
              tally() %>%
              mutate(p = n/sum(n)) %>%
              ungroup() 

Medidas_ordenado <- porcentaje %>% 
  filter(Valor == "I have applied this practice") %>% 
  arrange(n) %>% 
  pull(Medida)

porcentaje$Medida <- factor(porcentaje$Medida, levels = Medidas_ordenado)


conocidas_p <- porcentaje %>%
    pivot_wider(id_cols = Medida,
              names_from = Valor,
              values_from = p) %>%
  mutate(Medida = str_to_lower(Medida)) %>%
  arrange(desc(`I have applied this practice`), desc(`I know this practice but I have never applied it`)) %>%
  rowwise() %>%
  mutate(conocida = sum(`I know this practice but I have never applied it`, `I have applied this practice`)*100)

conocidas_n <- porcentaje %>%
    pivot_wider(id_cols = Medida,
              names_from = Valor,
              values_from = n) %>%
  mutate(Medida = str_to_lower(Medida)) %>%
  arrange(desc(`I have applied this practice`), desc(`I know this practice but I have never applied it`)) %>%
  rowwise() %>%
  mutate(conocida = sum(`I know this practice but I have never applied it`, `I have applied this practice`))

desconocidas_p <- conocidas_p %>%
  arrange(desc(`I don't know this practice`))

desconocidas_n <- conocidas_n %>%
  arrange(desc(`I don't know this practice`))



plot.height = n_distinct(porcentaje$Medida)*0.5+1

```



A total of `r nrow(publicacion)` participants (`r sprintf('%.1f', nrow(publicacion)/nrow(inscripciones)*100)`%) answered a question regarding the practice of openly publishing various products of the research process. Nevertheless, not every one completed every item:

```{r}

publicacion2 %>%
  count(Medida) %>%
  knitr::kable(col.names = c("Open practice", "N responses"),
               caption = "Table 1. Number of participants that answered regarding each open product.")

```


We can observed that:

- The most well-known practice was `r conocidas_p$Medida[1]` (`r sprintf('%.1f', conocidas_p[1,3]*100)`% - n = `r conocidas_n[1,3]` - of the participants mentioned knowing this practice, while `r sprintf('%.1f', conocidas_p[1,4]*100)`% - n = `r conocidas_n[1,4]` - had already applied it).

- The least known practices were `r str_c(desconocidas_p$Medida[1:3], collapse = ", ")` with `r str_c(sprintf('%.1f', (desconocidas_p[[2]][1:3]*100)), "% (n = ", desconocidas_n[[2]][1:3], ")", collapse = ", ")` respectively.


-   Each of the listed practices were known by at least `r sprintf('%.1f', min(conocidas_p$conocida))`% of the participants (n = `r min(conocidas_n$conocida)`), and used by more than a `r sprintf('%.1f', min(conocidas_p[4])*100)`% (n = `r min(conocidas_n[4])`).

#### Figure 12. Participants' knowledge regarding various Open Science practices.  
```{r, fig.width=7, fig.height=plot.height}
porcentaje %>%
  ggplot(aes(x = Medida, y = p, fill = Valor)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(y = "", x = "", fill = "") +
  theme_minimal(base_size = base_size) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(labels = function(x) str_wrap(x, width =25), values = c("#f67a21ff", "#00506fff", "#c63838ff")) +
  geom_text(aes(y = p, label = scales::percent(p, accuracy = 0.1)),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, color = "white") +
  coord_flip() +
  theme(    legend.position = "bottom",
    legend.justification = c(0, 0),  # Justify the legend to the left
    legend.box.just = "left",
    legend.margin = margin(t = 0, r = 0, b = 0, l = -200),  # Adjust left margin to move the legend leftwards
    panel.grid = element_blank())



```



```{r}
importancia <- investigacion %>% 
  select(importante_preregistrar:importante_resultados2) %>%
  mutate_all(~ ifelse(str_detect(.,"No sé"), NA, .))

importancia <- importancia[rowSums(is.na(importancia)) != ncol(importancia), ]

```

### How important do they think open practices are? (n = `r nrow(importancia)`, `r round(nrow(importancia)/nrow(inscripciones)*100,1)`%)

```{r}


importancia2 <- importancia %>% 
  pivot_longer(cols = everything(),
               names_to = "Medida",
               values_to = "Valor") %>% 
  mutate(Medida = fct_recode(Medida, `Pre-registrar proyectos` = "importante_preregistrar",
`Publicar instrumentos` = "importante_instrumentos",
`Publicar datos o código` = "importante_datosocodigo",
`Publicar resultados en revistas` = "importante_resultados1",
`Publicar resultados en medios alternativos` = "importante_resultados2")) %>% 
  mutate(Valor = ifelse(Valor %in% c("Poco importante", "Nada importante"), "Poco o no importante", Valor),
         Valor = fct_relevel(Valor, "Poco o no importante",
                             "Neutral",
                             "Importante",
                             "Muy importante")) %>%
  filter(!is.na(Valor))

porcentaje <- importancia2 %>% 
              group_by(Medida, Valor) %>%
              tally() %>%
              mutate(p = n/sum(n)) %>%
              ungroup() %>%
  mutate(etiqueta = ifelse(p < .059, NA, round(p,2)))

Medidas_ordenado <- porcentaje %>% 
  filter(Valor == "Muy importante") %>% 
  arrange(n) %>% 
  pull(Medida) 

porcentaje$Medida <- factor(porcentaje$Medida, levels = Medidas_ordenado)


porcentaje <- porcentaje %>% 
  mutate(Valor = fct_recode(Valor, "Very important" = "Muy importante",
                            "Important" ="Importante",
                            "Neutral" = "Neutral",
                            "Less or not important" = "Poco o no importante"),
         Medida = fct_recode(Medida,   "Pre-register projects" = "Pre-registrar proyectos",
  "Publish instruments" = "Publicar instrumentos",
  "Publish data or code" = "Publicar datos o código",
  "Disseminate results in alternative media" = "Publicar resultados en medios alternativos",
  "Publish results in Open Access journals" = "Publicar resultados en revistas"))

# Práctica que más personas consideran muy importante
very_important <- porcentaje %>%
  filter(Valor == "Very important") %>%
  mutate(p = p*100) %>%
  arrange(desc(p)) %>%
  slice(1)

# Práctica que más personas consideran poco o no importante
not_important <- porcentaje %>%
  filter(Valor == "Less or not important") %>%
  mutate(p = p*100) %>%
  arrange(desc(p)) %>%
  slice(1)


plot.height = n_distinct(porcentaje$Medida)*0.5+1


```

The practice identified as very important most often was "`r very_important$Medida`" (`r sprintf('%.1f', very_important$p)`%, n = `r very_important$n`), while the practice that more people considered of little or no importance was "`r not_important$Medida`" (`r sprintf('%.1f', not_important$p)`%, n = `r not_important$n`).

#### Figure 13. Levels of importance assigned to various Open Science practices by participants. 

```{r, fig.width=7, fig.height=plot.height}

porcentaje %>%  
  ggplot(aes(x = p, y = Medida, fill = Valor)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(y = "", x = "", fill = "") +
  scale_fill_manual(labels = function(x) str_wrap(x, width = 20), values = c("#5cb3c1", "#f67a21ff", "#00506fff", "#c63838ff")) +
  theme_minimal(base_size = base_size) +
  scale_x_continuous(labels = scales::percent) +
  geom_text(aes(x = etiqueta, label = scales::percent(etiqueta, accuracy = 0.1)),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, color = "white") +
  theme(
    legend.position = "bottom",
    legend.justification = c(0, 0),  # Justify the legend to the left
    legend.box.just = "left",
    legend.margin = margin(t = 0, r = 0, b = 0, l = -200),  # Adjust left margin to move the legend leftwards
    panel.grid = element_blank()
  )
  



```






```{r}
incentivar <- investigacion %>%
  filter(!is.na(animas)) %>%
  select(animas)

incentivar2 <- incentivar %>%
  mutate(animas = str_replace(animas, pattern = "difundir resultados en medios como blogs, publicaciones en redes sociales, o podcasts, entre otros?", replacement = "difundir resultados en medios alternativos")) %>% 
  separate(col = animas, 
           into= c("animas1", "animas2", "animas3", "animas4", "animas5", "animas6"), sep=",") %>%
  pivot_longer(
    cols = animas1:animas6,
    names_to = "col",
    values_to = "animas"
  ) %>%
  mutate(animas = str_trim(animas, side = "both"),
         animas = str_to_sentence(animas))%>%
  filter(!is.na(animas)) %>%
  group_by(animas) %>%
  summarise(cantidad = n()) %>% 
  arrange(desc(cantidad)) %>% 
  mutate(por = cantidad/nrow(incentivar)*100)


incentivar2 <- incentivar2 %>%
  mutate(animas = recode(animas, "Difundir resultados en medios alternativos?" = "Share results in alternative media",
  "Publicar resultados en revistas de acceso abierto?" = "Publish results in open access journals",
  "Publicar datos o código en repositorios abiertos?" = "Publish data or code in open repositories",
  "Publicar los instrumentos usados para recolectar datos?" = "Publish the instruments used to collect data",
  "Pre-registrar sus proyectos?" = "Pre-register their projects"))

plot.height = n_distinct(incentivar2$animas)*0.5+0.5

```

### Do they encourage other people to go open? (n = `r nrow(incentivar)`, `r round(nrow(incentivar)/nrow(inscripciones)*100,1)`%)

```{r}

# Filtrar los que tienen más del 50% de respuestas
over_50 <- incentivar2 %>% filter(por > 50) %>%
  mutate(animas = str_to_lower(animas))


# Filtrar el que tiene el menor porcentaje
least_incentivated <- incentivar2 %>% 
  filter(por == min(por)) %>%
  mutate(animas = str_to_lower(animas))

# Generar el texto automáticamente
result_text <- glue(
  "Out of the total of people who answered this question, more than 50% encourages their colleagues and thesis students to {paste(over_50$animas, collapse = ', ')} ({paste(round(over_50$por, 1), '% - n = ', over_50$cantidad, sep = '', collapse = ', ')}%, respectively). In agreement with answers to the previous question, the practice they reported to incentivize the least was {least_incentivated$animas} ({round(least_incentivated$por, 1)}%, n = {least_incentivated$cantidad})."
)




```


We asked participants if they encourage other professionals, researchers, their coworkers, and/or their thesis students or advisors to implement various open practices. 
`r result_text`

#### Figure 14. Open practices the participants encourage others to implement.

```{r, fig.width=7, fig.height=plot.height}
incentivar2 %>%
  ggplot(aes(fct_reorder(animas, por), por)) +
  geom_bar(stat = "identity",fill="#c83737") +
  ggtitle("")+
  coord_flip() +
  scale_y_continuous(expand = c(0,0)) +
  theme_minimal(base_size = base_size) +
  labs(caption = "More than one option can be selected.", y = "Percentage (%)", x = "")

```

The total percentage exceeds 100% because participants were asked to choose all the practices they encourage others to implement.


## Same metrics after the cohort finished

```{r}
publicacion <- encuesta_final %>% 
  select(conoces_proyecto:conoces_resultados2) %>% 
   mutate_all(~ ifelse(. == "No aplica/No sé/Prefiero no contestar", NA, .))

publicacion <- publicacion[rowSums(is.na(publicacion)) != ncol(publicacion), ]


```

The final survey was completed by `r nrow(publicacion)` participants (`r sprintf('%.1f', nrow(publicacion)/nrow(inscripciones)*100)`% of the people that signed up and `r sprintf('%.1f', nrow(publicacion)/n_distinct(zoom_tiempo$Nombre_corregido)*100)`% of the people that attendeed the zoom meetings).


### What is their knowledge about open publishing after having completed the Cohort training? (n = `r nrow(publicacion)`, `r round(nrow(publicacion)/nrow(inscripciones)*100,1)`%)


```{r}

## Español
# publicacion2 <- publicacion %>% 
#   pivot_longer(cols = everything(),
#                names_to = "Medida",
#                values_to = "Valor") %>% 
#   mutate(Medida = fct_recode(Medida, `Proyecto de investigación` = "conoces_proyecto",
# `Plan de gestión de datos` = "conoces_pgd",
# `Hipótesis y plan de análisis` = "conoces_hipotesis",
# `Datos recolectados` = "conoces_datos",
# `Código` = "conoces_codigo",
# `Instrumentos` = "conoces_instrumentos",
# `Resultados en revistas` = "conoces_resultados1",
# `Resultados en medios alternativos` = "conoces_resultados2")) %>% 
#   mutate(Valor = fct_relevel(Valor, "No conozco esta práctica",
#                              "Conozco la práctica pero nunca la he aplicado",
#                              "He aplicado la práctica")) %>% 
#   filter(!is.na(Valor))

# Inglés
publicacion2 <- publicacion %>% 
  pivot_longer(cols = everything(),
               names_to = "Medida",
               values_to = "Valor") %>% 
  mutate(Medida = fct_recode(Medida, `Pre-register a research project` = "conoces_proyecto",
`Data management plan` = "conoces_pgd",
`Hypothesis and analysis plan` = "conoces_hipotesis",
`Data collected` = "conoces_datos",
`Code` = "conoces_codigo",
`Instruments` = "conoces_instrumentos",
`Results in Open Access journals` = "conoces_resultados1",
`Results in alternative media` = "conoces_resultados2")) %>% 
  mutate(Valor = fct_relevel(Valor, "No conozco esta práctica",
                             "Conozco la práctica pero nunca la he aplicado",
                             "He aplicado la práctica"),
         Valor = fct_recode(Valor, 
                            "I don't know this practice" = "No conozco esta práctica",
  "I know this practice but I have never applied it" = "Conozco la práctica pero nunca la he aplicado",
  "I have applied this practice" = "He aplicado la práctica")) %>% 
  filter(!is.na(Valor))




porcentaje <- publicacion2 %>% 
              group_by(Medida, Valor) %>%
              tally() %>%
              mutate(p = n/sum(n)) %>%
              ungroup() 

Medidas_ordenado <- porcentaje %>% 
  filter(Valor == "I have applied this practice") %>% 
  arrange(n) %>% 
  pull(Medida)

porcentaje$Medida <- factor(porcentaje$Medida, levels = Medidas_ordenado)


conocidas_p <- porcentaje %>%
    pivot_wider(id_cols = Medida,
              names_from = Valor,
              values_from = p) %>%
  mutate(Medida = str_to_lower(Medida)) %>%
  arrange(desc(`I have applied this practice`), desc(`I know this practice but I have never applied it`)) %>%
  rowwise() %>%
  mutate(conocida = sum(`I know this practice but I have never applied it`, `I have applied this practice`)*100)

conocidas_n <- porcentaje %>%
    pivot_wider(id_cols = Medida,
              names_from = Valor,
              values_from = n) %>%
  mutate(Medida = str_to_lower(Medida)) %>%
  arrange(desc(`I have applied this practice`), desc(`I know this practice but I have never applied it`)) %>%
  rowwise() %>%
  mutate(conocida = sum(`I know this practice but I have never applied it`, `I have applied this practice`))

desconocidas_p <- conocidas_p %>%
  arrange(desc(`I don't know this practice`))

desconocidas_n <- conocidas_n %>%
  arrange(desc(`I don't know this practice`))



plot.height = n_distinct(porcentaje$Medida)*0.5+1

```

A total of `r nrow(publicacion)` participants (`r sprintf('%.1f', nrow(publicacion)/nrow(inscripciones)*100)`%) answered a question regarding the practice of openly publishing various products of the research process. Nevertheless, not every one completed every item:

```{r}

publicacion2 %>%
  count(Medida) %>%
  knitr::kable(col.names = c("Open practice", "N responses"),
               caption = "Table 1. Number of participants that answered regarding each open product.")

```


We can observed that:

- The most well-known practice was `r conocidas_p$Medida[1]` (`r sprintf('%.1f', conocidas_p[1,3]*100)`% - n = `r conocidas_n[1,3]` - of the participants mentioned knowing this practice, while `r sprintf('%.1f', conocidas_p[1,4]*100)`% - n = `r conocidas_n[1,4]` - had already applied it).

- The least known practices were `r str_c(desconocidas_p$Medida[1:3], collapse = ", ")` with `r str_c(sprintf('%.1f', (desconocidas_p[[2]][1:3]*100)), "% (n = ", desconocidas_n[[2]][1:3], ")", collapse = ", ")` respectively.


-   Each of the listed practices were known by at least `r sprintf('%.1f', min(conocidas_p$conocida))`% of the participants (n = `r min(conocidas_n$conocida)`), and used by more than a `r sprintf('%.1f', min(conocidas_p[4])*100)`% (n = `r min(conocidas_n[4])`).

#### Figure 15. Participants' knowledge regarding various Open Science practices after completing the Cohort training.

```{r, fig.width=7, fig.height=plot.height}
porcentaje %>%
  ggplot(aes(x = Medida, y = p, fill = Valor)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(y = "", x = "", fill = "") +
  theme_minimal(base_size = base_size) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(labels = function(x) str_wrap(x, width =25), values = c("#f67a21ff", "#00506fff", "#c63838ff")) +
  geom_text(aes(y = p, label = scales::percent(p, accuracy = 0.1)),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, color = "white") +
  coord_flip() +
  theme(    legend.position = "bottom",
    legend.justification = c(0, 0),  # Justify the legend to the left
    legend.box.just = "left",
    legend.margin = margin(t = 0, r = 0, b = 0, l = -200),  # Adjust left margin to move the legend leftwards
    panel.grid = element_blank())



```



```{r}
importancia <- encuesta_final %>% 
  select(importante_preregistrar:importante_resultados2) %>%
  mutate_all(~ ifelse(str_detect(.,"No sé"), NA, .))

importancia <- importancia[rowSums(is.na(importancia)) != ncol(importancia), ]

```

### How important do they think open practices are after having completed the Cohort training? (n = `r nrow(importancia)`, `r round(nrow(importancia)/nrow(inscripciones)*100,1)`%)

```{r}


importancia2 <- importancia %>% 
  pivot_longer(cols = everything(),
               names_to = "Medida",
               values_to = "Valor") %>% 
  mutate(Medida = fct_recode(Medida, `Pre-registrar proyectos` = "importante_preregistrar",
`Publicar instrumentos` = "importante_instrumentos",
`Publicar datos o código` = "importante_datosocodigo",
`Publicar resultados en revistas` = "importante_resultados1",
`Publicar resultados en medios alternativos` = "importante_resultados2")) %>% 
  mutate(Valor = ifelse(Valor %in% c("Poco importante", "Nada importante"), "Poco o no importante", Valor),
         Valor = fct_relevel(Valor, "Poco o no importante",
                             "Neutral",
                             "Importante",
                             "Muy importante")) %>%
  filter(!is.na(Valor))

porcentaje <- importancia2 %>% 
              group_by(Medida, Valor) %>%
              tally() %>%
              mutate(p = n/sum(n)) %>%
              ungroup() %>%
  mutate(etiqueta = ifelse(p < .059, NA, round(p,2)))

Medidas_ordenado <- porcentaje %>% 
  filter(Valor == "Muy importante") %>% 
  arrange(n) %>% 
  pull(Medida) 

porcentaje$Medida <- factor(porcentaje$Medida, levels = Medidas_ordenado)


porcentaje <- porcentaje %>% 
  mutate(Valor = fct_recode(Valor, "Very important" = "Muy importante",
                            "Important" ="Importante",
                            "Neutral" = "Neutral",
                            "Less or not important" = "Poco o no importante"),
         Medida = fct_recode(Medida,   "Pre-register projects" = "Pre-registrar proyectos",
  "Publish instruments" = "Publicar instrumentos",
  "Publish data or code" = "Publicar datos o código",
  "Disseminate results in alternative media" = "Publicar resultados en medios alternativos",
  "Publish results in Open Access journals" = "Publicar resultados en revistas"))

# Práctica que más personas consideran muy importante
very_important <- porcentaje %>%
  filter(Valor == "Very important") %>%
  mutate(p = p*100) %>%
  arrange(desc(p)) %>%
  slice(1)

# Práctica que más personas consideran poco o no importante
not_important <- porcentaje %>%
  filter(Valor == "Less or not important") %>%
  mutate(p = p*100) %>%
  arrange(desc(p)) %>%
  slice(1)


plot.height = n_distinct(porcentaje$Medida)*0.5+1


```

The practice identified as very important most often was "`r very_important$Medida`" (`r sprintf('%.1f', very_important$p)`%, n = `r very_important$n`), while the practice that more people considered of little or no importance was "`r not_important$Medida`" (`r sprintf('%.1f', not_important$p)`%, n = `r not_important$n`).

#### Figure 16. Levels of importance assigned to various Open Science practices by participants. 

```{r, fig.width=7, fig.height=plot.height}

porcentaje %>%  
  ggplot(aes(x = p, y = Medida, fill = Valor)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(y = "", x = "", fill = "") +
  scale_fill_manual(labels = function(x) str_wrap(x, width = 20), values = c("#5cb3c1", "#f67a21ff", "#00506fff", "#c63838ff")) +
  theme_minimal(base_size = base_size) +
  scale_x_continuous(labels = scales::percent) +
  geom_text(aes(x = etiqueta, label = scales::percent(etiqueta, accuracy = 0.1)),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, color = "white") +
  theme(
    legend.position = "bottom",
    legend.justification = c(0, 0),  # Justify the legend to the left
    legend.box.just = "left",
    legend.margin = margin(t = 0, r = 0, b = 0, l = -200),  # Adjust left margin to move the legend leftwards
    panel.grid = element_blank()
  )
  



```



```{r}
incentivar <- encuesta_final %>%
  filter(!is.na(animas)) %>%
  select(animas)

incentivar2 <- incentivar %>%
  mutate(animas = str_replace(animas, pattern = "difundir resultados en medios como blogs, publicaciones en redes sociales, o podcasts, entre otros?", replacement = "difundir resultados en medios alternativos")) %>% 
  separate(col = animas, 
           into= c("animas1", "animas2", "animas3", "animas4", "animas5", "animas6"), sep=",") %>%
  pivot_longer(
    cols = animas1:animas6,
    names_to = "col",
    values_to = "animas"
  ) %>%
  mutate(animas = str_trim(animas, side = "both"),
         animas = str_to_sentence(animas))%>%
  filter(!is.na(animas)) %>%
  group_by(animas) %>%
  summarise(cantidad = n()) %>% 
  arrange(desc(cantidad)) %>% 
  mutate(por = cantidad/nrow(incentivar)*100)


incentivar2 <- incentivar2 %>%
  mutate(animas = recode(animas, "Difundir resultados en medios alternativos?" = "Share results in alternative media",
  "Publicar resultados en revistas de acceso abierto?" = "Publish results in open access journals",
  "Publicar datos o código en repositorios abiertos?" = "Publish data or code in open repositories",
  "Publicar los instrumentos usados para recolectar datos?" = "Publish the instruments used to collect data",
  "Pre-registrar sus proyectos?" = "Pre-register their projects"))

plot.height = n_distinct(incentivar2$animas)*0.5+0.5

```

### Will they encourage other people to go open after having completed the Cohort training? (n = `r nrow(incentivar)`, `r round(nrow(incentivar)/nrow(inscripciones)*100,1)`%)

```{r}

# Filtrar los que tienen más del 50% de respuestas
over_50 <- incentivar2 %>% filter(por > 50) %>%
  mutate(animas = str_to_lower(animas))


# Filtrar el que tiene el menor porcentaje
least_incentivated <- incentivar2 %>% 
  filter(por == min(por)) %>%
  mutate(animas = str_to_lower(animas))

# Generar el texto automáticamente
result_text <- glue(
  "Out of the total of people who answered this question, more than 50% states that they will encourage their colleagues and thesis students to {paste(over_50$animas, collapse = ', ')} ({paste(round(over_50$por, 1), '% - n = ', over_50$cantidad, sep = '', collapse = ', ')}%, respectively). In agreement with answers to the previous question, the practice they they consider they will incentivize the least is {least_incentivated$animas} ({round(least_incentivated$por, 1)}%, n = {least_incentivated$cantidad})."
)




```


We asked participants if after having participated in the Cohort they will encourage other professionals, researchers, their coworkers, and/or their thesis students or advisors to implement various open practices. 
`r result_text`

#### Figure 17. Open practices the participants will encourage others to implement.

```{r, fig.width=7, fig.height=plot.height}
incentivar2 %>%
  ggplot(aes(fct_reorder(animas, por), por)) +
  geom_bar(stat = "identity",fill="#c83737") +
  ggtitle("")+
  coord_flip() +
  scale_y_continuous(expand = c(0,0)) +
  theme_minimal(base_size = base_size) +
  labs(caption = "More than one option can be selected.", y = "Percentage (%)", x = "")

```

The total percentage exceeds 100% because participants were asked to choose all the practices they encourage others to implement.




## Zoom attendance



Out of the `r nrow(inscripciones)` people who signed up for the cohort, `r n_distinct(zoom_tiempo$Nombre_corregido)` (`r round(n_distinct(zoom_tiempo$Nombre_corregido)/nrow(inscripciones)*100, 1)`%) attended the Zoom meetings at least once over the 6 synchronous sessions.

 

```{r}

zoom_final %>% 
  mutate(total = nrow(inscripciones),
         por = round(n.x/total*100,1)) %>% 
  select(Encuentro, n.x, por) %>% 
  mutate(Encuentro = str_remove(Encuentro, pattern = "Encuentro ")) %>% 
  knitr::kable(col.names = c("Meeting", "N", "%"),
               caption = "Table 1. Number of participants that attendeed each meeting.") %>%
add_footnote("Note. N = number of attendees; % = percentage of attendees over total of participants." )

asistencia <- IC_funcion(data = zoom_final, variable = "n.x")

```

The mean number of attendees per meeting was `r sprintf('%.1f',asistencia$media)` (SD = `r sprintf('%.1f', asistencia$DE)`, CI 95% = `r asistencia$IC`).

```{r}

zoom_final %>% 
  select(-n.x) %>% 
  mutate(final = round(final, 1)) %>%
  mutate(Encuentro = str_remove(Encuentro, pattern = "Encuentro ")) %>% 
  knitr::kable(col.names = c("Meeting", "Remained until the end", "%"),
               caption = "Table 2. Number of participants that attendeed each meeting.") %>%
add_footnote("Note. N = number of attendees that remained until the end of the meeting; % = percentage over total of attendees." )

Duracion <- IC_funcion(data = zoom_tiempo, variable = "Duracion")

# Calcular duración media en minutos
mean_minutes <- round(Duracion$media, 0)

# Convertir minutos en horas y minutos
hours <- mean_minutes %/% 60
minutes <- round(mean_minutes %% 60, 0)

# Calcular ic_inf y ic_sup en horas y minutos
ic_sup_hours <- Duracion$ICs %/% 60
ic_sup_minutes <- round(Duracion$ICs %% 60, 0)

ic_inf_hours <- Duracion$ICi %/% 60
ic_inf_minutes <- round(Duracion$ICi %% 60, 0)

por_hora_total <- (mean_minutes/180)*100
por_hora_total <- sprintf('%.1f', por_hora_total)

```

A mean of `r round(mean(zoom_final$final),1)`% of the attendees remained connected until the end of the meetings. Over the 6 meetings, the mean connection time was of `r paste(hours, "hours and", minutes, "minutes")`, (SD = `r round(sd(zoom_tiempo$Duracion, na.rm = TRUE),0)` minutes, CI 95% = `r paste(ic_inf_hours, "hours and", ic_inf_minutes, "minutes")`, `r paste(ic_sup_hours, "hours and", ic_sup_minutes, "minutes")`), a `r por_hora_total`% of the mean connection time that the sessions lasted on average 3 hours.

#### Figure 18. Number of participants that attendeed each meeting and remained until the end of the session. 

```{r}

zoom_final %>% 
  mutate(Encuentro = str_remove(Encuentro, pattern = "Encuentro ")) %>% 
  pivot_longer(cols = n.x:n.y,
               names_to = "asistente",
               values_to = "n") %>% 
  ggplot(aes(x = as.numeric(Encuentro), y = n, color = asistente, linetype = asistente)) +
  geom_line(size = 1) +
  labs(x = "# Meeting", y = "Number of participants",
       title = "Zoom attendance") +
  ylim(0, 60) +
  theme_minimal(base_size = base_size) +
  scale_color_manual(values = c("n.y" = "#00506F", "n.x" = "#C83737"), labels = c("Attended the meeting", "Remained connected until the end")) +
  scale_linetype_manual(values = c("n.y" = "dashed", "n.x" = "solid"), labels = c("Attended the meeting", "Remained connected until the end")) +
  guides(color = guide_legend(title = NULL), linetype = guide_legend(title = NULL)) +
  theme(legend.position = "bottom")


```


## NASA TOPS certification 

```{r}


certificaciones <- evaluaciones %>%
  filter(forms_completados == 5) %>%
  count()

por_resp <- nrow(evaluaciones)/nrow(inscripciones)*100
por_cert <- certificaciones$n/nrow(inscripciones)*100

```

Out of the `r nrow(inscripciones)` people who signed up, `r nrow(evaluaciones)` (`r sprintf('%.1f', por_resp)`%) completed at least one of the assessment forms and `r certificaciones$n` (`r sprintf('%.1f', por_cert)`%) completed all required assessments to get certified by NASA.

#### Figure 19. Number of participants that approved the assessment form. 

```{r}

evaluaciones %>%
  select(E1:E5) %>%
  pivot_longer(cols = E1:E5,
               names_to = "Encuentro",
               values_to = "n", names_prefix = "E") %>% 
  group_by(Encuentro) %>%
  summarise(n = sum(n)) %>%
  ggplot(aes(x = as.numeric(Encuentro), y = n)) +
  geom_line(size = 1, color = "#C83737") +
  labs(x = "# Meeting", y = "Number of participants",
       title = "Number of assessments completed", color = "") +
  ylim(0,70) +
 theme_minimal(base_size = base_size) 

```


## Joined slack


Out of the `r nrow(inscripciones)` people who signed up for this first cohort, `r slack$n` (`r round(slack$n/nrow(inscripciones)*100,1)`%) joined our Slack community! Welcome new members!

## Net Promoter Score

```{r}

# Remove NAs
responses <- encuesta_final$recomendacion[!is.na(encuesta_final$recomendacion)]

# Categorize respondents
promoters <- sum(responses >= 9)
passives <- sum(responses >= 7 & responses <= 8)
detractors <- sum(responses <= 6)

# Calculate percentages
total_respondents <- length(responses)
percent_promoters <- promoters / total_respondents * 100
percent_detractors <- detractors / total_respondents * 100

# Calculate NPS
nps <- percent_promoters - percent_detractors




```


The Net Promoter Score (NPS) (Reichheld, 2003) is a widely used metric that measures customer loyalty and satisfaction. It is calculated by asking respondents a single question: "How likely are you to recommend this course to a friend or colleague?" Responses are given on a scale from 0 to 10, where 0 means "Not at all likely" and 10 means "Extremely likely". Based on their responses, customers are categorized into three groups: Promoters (scores of 9-10), Passives (scores of 7-8), and Detractors (scores of 0-6). The NPS is determined by subtracting the percentage of Detractors from the percentage of Promoters. The resulting score can range from -100 to +100, with a higher score indicating greater customer satisfaction and loyalty.

[Suggested cutoff values](https://www.qualtrics.com/en-gb/experience-management/customer/good-net-promoter-score/?rid=ip&prevsite=en&newsite=uk&geo=GB&geomatch=uk) are:

- Above 0 is good,
- Above 20 is favorable,
- Above 50 is excellent, and
- Above 80 is world class.

Based on the `r length(responses)` participants that responded to this question in the final survey (`r sprintf('%.1f', length(responses)/n_distinct(zoom_tiempo$Nombre_corregido)*100)`% of attendees), an **NPS of `r sprintf('%.1f', nps)`%** was obtained, which is considered excellent. This score reflects a high level of satisfaction and loyalty among the respondents, indicating that they are likely to recommend our course to others.

Reichheld, F. F. (2003). The one number you need to grow. Harvard business review, 81(12), 46-55.


